---
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8,
                      echo=FALSE, warning=FALSE, message=FALSE)
```


```{r, echo=FALSE}
#Author: Mark Dane, copyright 2015
source("MEPLINCSFunctions.R")
supportingPlots <- FALSE
runRUV2S <-TRUE
runRep <- TRUE


```



```{r setup}
library("ggplot2")
library("reshape2")
library("data.table")
library("limma")
library("MEMA")
library("grid")
library("knitr")
library("gplots")
library("RColorBrewer")
library("RUVnormalize")

#Setup colors for Barcode and text in all heatmaps
selDark2 <- colorRampPalette(brewer.pal(8,"Dark2"))
plateCol = selDark2(8)

#ToDo Make changes for new folder sturcture
l1 <- fread(x[["inputFileName"]], showProgress = FALSE)

#Keep the control wells separate
l1$Ligand[l1$Ligand=="FBS"] <- paste(l1$Ligand[l1$Ligand=="FBS"],"C",as.numeric(factor(l1$Barcode[l1$Ligand=="FBS"])),sep="")
l1$MEP <- paste(l1$ECMp,l1$Ligand,sep="_")
setkey(l1, "ECMp")
l1 <- l1[!"fiducial"]
l1 <- l1[!"blank"]

#Logit transform DNA2NProportion
#logit(p) = log[p/(1-p)]
DNA2NImpute <- l1$Nuclei_PA_Cycle_DNA2NProportion
DNA2NImpute[DNA2NImpute==0] <- .01
DNA2NImpute[DNA2NImpute==1] <- .99
l1$Nuclei_PA_Cycle_DNA2NProportionLogit <- log2(DNA2NImpute/(1-DNA2NImpute))
l1$Cytoplasm_PA_Intensity_MedianIntensity_LinRatio <- log2((l1$Cytoplasm_CP_Intensity_MedianIntensity_KRT19+.001)/(l1$Cytoplasm_CP_Intensity_MedianIntensity_KRT5+.001))

#Create spot level dataset
l3 <- l1[,list(Ligand=unique(Ligand), LigandAnnotID=unique(LigandAnnotID),
               ECMp=unique(ECMp), ECMpAnnotID=unique(ECMpAnnotID),
               MEP=unique(MEP),
               Row=unique(Row), Column = unique(Column), Block=unique(Block), ID=unique(ID),
               ArrayRow=unique(ArrayRow), ArrayColumn=unique(ArrayColumn),
               LinRatiomel = median(Cytoplasm_PA_Intensity_MedianIntensity_LinRatio, na.rm=TRUE),
               DNA2Nmel = median(Nuclei_PA_Cycle_DNA2NProportionLogit, na.rm=TRUE),
               SCCmel = as.numeric(median(Spot_PA_SpotCellCount, na.rm=TRUE))
           ),
         by="Barcode,Well,Spot"]


l1MEPs <- l1[,list(mel = median(Cytoplasm_PA_Intensity_MedianIntensity_LinRatio, na.rm = TRUE),
                   Mel = mad(Cytoplasm_PA_Intensity_MedianIntensity_LinRatio, na.rm= TRUE)),
             by="Barcode,Well,Ligand,ECMp,MEP"]

#l1MEPS is the data.table of the MEP level, raw, log2-transformed LinRatio responses
l1MEPs <- addMarginValues(l1MEPs,"mel","Mel")


#calculate the overall median and MAD for the staining set
m.. <- median(l1MEPs$mel, na.rm = TRUE)
M.. <- median(l1MEPs$Mel, na.rm = TRUE)
```

```{r ReplicateRUV, eval=runRep}
#Unit of study is array so there are 64 'samples'of 594 spots with 8 replicates
#Order the spot level data by plate, well, spot number
#Create a spot assignment that works with rotated B row arrays
l3$RSpot <- l3$Spot
l3$RSpot[grepl("B",l3$Well)] <- 701-l3$RSpot[grepl("B",l3$Well)]
#Add in ligand and ECMp names so they will carry through the normalization
l3$BWL <- paste(l3$Barcode,l3$Well,l3$Ligand,sep="_")
l3$SE <- paste(l3$RSpot,l3$ECMp,sep="_")
l3$BW <- paste(l3$Barcode,l3$Well,sep="_")
#Cast to get mel values by barcode_well rows and spot columns
#Coerce missing values to have near 0 proliferation signals
l3LinRatioc <- dcast(l3, BWL~SE, value.var="LinRatiomel",fill = log2(.01/(1-.01)))
#Remove the BWL column and use it as rownames in the matrix
l3LinRatiom <- l3LinRatioc[,grep("BWL",colnames(l3LinRatioc), value=TRUE, invert=TRUE), with=FALSE]
#YLinRatio is a numeric matrix of the raw logit transformed LinRatio responses 
#of each spot extracted from the l3 dataset
YLinRatio <- matrix(unlist(l3LinRatiom), nrow=nrow(l3LinRatiom), dimnames=list(l3LinRatioc$BWL, colnames(l3LinRatiom)))

#Set up replicate index matrix with A03 wells as replicates
replicateWells <- which(grepl("A03",rownames(YLinRatio)))
scIdx <- matrix(-1, 57, 8)
c <- 1
scIdx[1,] <- replicateWells

for (r in 1:nrow(YLinRatio)){
  if(r %in% replicateWells)
    next
  c <- c+1
  scIdx[c,1] <- r
}

```
---
title: "MEP-LINCS `r x[["CellLine"]]` `r x[["StainingSet"]]` `r x[["Signal"]]` `r x[["Method"]]` Normalization"
date: "`r Sys.Date()`"
---


###Lineage Ratio Signal
This analysis focuses on the ratio of the median cytoplasmic intensity values of KRT19/KRT5 at each spot. The lineage ratios in this dataset have been log2 transformed into the range [-10, 10].  


```{r, eval=TRUE, fig.width=4.5, fig.height=4}


titleExpression <- expression(LinRatioMedians~(m["e,l"]))
plotLEHmap(l1MEPs, fill="mel", titleExpression)

titleExpression <- expression(LinRatio^{"+"}~MADs~(M["e,l"]))
plotLEHmap(l1MEPs, fill="Mel", titleExpression)


```

```{r}

nYLinRatioList <- lapply(1:7, function(k, YLinRatio, cIdx, scIdx, mValue, MValue){
    #Normalize the spot level data then median summarize
    nYLinRatiom <- medianNaiveReplicateRUV2(k, YLinRatio, cIdx, scIdx)
    setnames(nYLinRatiom, "Value", "mel")
    #Normalize the spot level data then MAD summarize
    nYLinRatioM <- madNaiveReplicateRUV2(k, YLinRatio, cIdx, scIdx)
    setnames(nYLinRatioM, "Value", "Mel")
    setkey(nYLinRatiom,Barcode,MEP,ECMp,Ligand)
    setkey(nYLinRatioM, Barcode,MEP,ECMp,Ligand)
    nYLinRatio <- merge(nYLinRatiom, nYLinRatioM)
    #Add margin values
    nYLinRatio <- addMarginValues(dt=nYLinRatio, mValue="mel",MValue="Mel")
    nYLinRatio$k <- k
    return(nYLinRatio)
}, YLinRatio=YLinRatio, cIdx=1:ncol(YLinRatio), scIdx=scIdx, mValue="mel", MValue="Mel")

#nYLinRatiok is the RUV2 normalized LinRatio responses at different k levels
#using the array as the unit of study
#All spots are identified as controls

nYLinRatioArrayk <- rbindlist(nYLinRatioList)

```

```{r, fig.width=4.5, fig.height=3}
groupBy <- "ECMp"
value <- "mel-m.l"
colourBy <- "Barcode"
titleExpression=expression(LinRatio^{"+"}~Centered~Medians~by~ECMp~(m["e,l"]*"-"*m[".,l"]))
p <- ggplot(l1MEPs, aes_string(x=groupBy, y=value))+
  geom_boxplot()+ 
  ggtitle(titleExpression)+
  xlab("")+ylab("")+
  #ylim(-7,7)+
  guides(fill=FALSE)+
  theme_grey(base_size = 12, base_family = "")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.6)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(.5)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.5)))
suppressWarnings(print(p))
```

###NaiveReplicateRUV2 with Array as Unit of Study

In the following plots naiveReplicateRUV-2 normalization is applied to the LinRatio signal of all spots with an array as the unit of study. The input data is a 64x694 matrix of logit transformed LinRatio+ values. The A03 FBS wells are designated as replicates. All spots are designated as negative controls. The faceted plots show various values of the nu and k parameters.

```{r}

dt <- nYLinRatioArrayk
p <- ggplot(dt,aes_string(x=groupBy, y=value))+
  geom_boxplot()+ 
  ggtitle(bquote(LinRatioMedians~(m["e,l"]*"-"*m[".,l"])~After~NaiveReplicateRUV2~Normalization~with ~Array~Unit~of~Study))+
  xlab("")+ylab("")+
  #ylim(-7,7)+
  guides(fill=FALSE)+
  theme_grey(base_size = 12, base_family = "")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.6)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.5)))+
  facet_grid(k~., labeller = labeller(k = label_both))
suppressWarnings(print(p))
```

```{r, fig.width=4.5, fig.height=3}
groupBy <- "Ligand"
value <- "mel-me."
colourBy <- "Barcode"
titleExpression=expression(LinRatio^{"+"}~Centered~Medians~by~Ligand~(m["e,l"]*"-"*m["e,."]))
p <- ggplot(l1MEPs, aes_string(x=groupBy, y=value, colour=colourBy))+
  geom_boxplot()+ 
  ggtitle(titleExpression)+
  xlab("")+ylab("")+
  #ylim(-7,7)+
  guides(fill=FALSE)+
  theme_grey(base_size = 12, base_family = "")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.6)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(.5)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.5)))
suppressWarnings(print(p))
```

```{r}

dt <- nYLinRatioArrayk
p <- ggplot(dt,aes_string(x=groupBy, y=value, colour=colourBy))+
  geom_boxplot()+ 
  ggtitle(bquote(LinRatioMedians~(m["e,l"]*"-"*m["e.,."])~After~NaiveReplicateRUV2~Normalization~with ~Array~Unit~of~Study))+
  xlab("")+ylab("")+
  #ylim(-7,7)+
  guides(fill=FALSE)+
  theme_grey(base_size = 12, base_family = "")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.6)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.5)))+
  facet_grid(k~., labeller = labeller(k = label_both))
suppressWarnings(print(p))
```


```{r, fig.width=5, fig.height=4}

fill <- "mel"
titleExpression <- expression(LinRatioMedians~(m["e,l"]))
dt <- l1MEPs
limits <- quantile(dt[[fill]], probs=c(.02, .98))

p <- ggplot(dt, aes_string(x="Ligand", y="ECMp", fill = fill))+
  geom_tile()+
  scale_fill_gradient(low="white",high="red",oob = scales::squish,limits=limits)+
  ggtitle(titleExpression)+
  xlab("")+ylab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.6)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(.5)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)),panel.grid.major = element_line(linetype = 'blank'),panel.background = element_rect(fill = "dimgray"))
suppressWarnings(print(p))
```

```{r, fig.width=4.5}
dt <- nYLinRatioArrayk

p <- ggplot(dt, aes_string(x="Ligand", y="ECMp", fill = fill))+
  geom_tile()+
  scale_fill_gradient(low="white",high="red",oob = scales::squish, limits=limits)+
  ggtitle((bquote(LinRatioMedians~(m["e,l"]*"-"*m["e.,."])~After~NaiveReplicateRUV2~Normalization~with ~Array~Unit~of~Study)))+
  facet_grid(k~., labeller = labeller(k = label_both))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.6)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(.5)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)),panel.grid.major = element_line(linetype = 'blank'),panel.background = element_rect(fill = "dimgray"))
suppressWarnings(print(p))

```
