---
output: html_document
---
---
title: "MEP-LINCS Source Plate Analysis"
date: "`r Sys.Date()`"
---

```{r global_options_Setup, include=FALSE}
#Author: Mark Dane, copyright 2015
knitr::opts_chunk$set(fig.width=12, fig.height=8,
                      echo=FALSE, warning=FALSE, message=FALSE)

#source("MEPLINCSFunctions.R")

library("ggplot2")
library("reshape2")
library("data.table")
library("limma")
library("MEMA")
library("grid")
library("knitr")
library("gplots")
library("RColorBrewer")
library("RUVnormalize")
library(plotly)

#Read in file based on name in dataframe
l1List <- lapply(x[["inputFileName"]], function(fn){
  DT <- fread(fn, showProgress = FALSE)
  DT <- DT[,grep("ECMp|Ligand|MEP|^ID$|Array|_SpotCellCount|Row|Column|Block|^Spot$|Barcode|^Well$|WellIndex",colnames(DT), value=TRUE), with=FALSE]
})

l1 <- rbindlist(l1List)
#Keep the control wells separate
l1$Ligand[l1$Ligand=="FBS"] <- paste(l1$Ligand[l1$Ligand=="FBS"],"C",as.numeric(factor(l1$Barcode[l1$Ligand=="FBS"])),sep="")
l1$MEP <- paste(l1$ECMp,l1$Ligand,sep="_")
l1 <- l1[, ECMSourcePlateSeries := as.factor(tstrsplit(x=Barcode,split="")[[7]])]
l1 <- l1[,PlatePrintOrder := as.numeric(paste0(tstrsplit(x=Barcode,split="")[[8]],
               tstrsplit(x=Barcode,split="")[[9]]))]
l1 <- l1[, ArrayPrintOrder := (PlatePrintOrder-1)*8+WellIndex]
#Delete the fiducial and blank spots
setkey(l1, "ECMp")
l1 <- l1[!"fiducial"]

#Create spot level dataset and 
#add Rel, the residual from subtracting the mel from each value

l3 <- l1[,list(Ligand=unique(Ligand), LigandAnnotID=unique(LigandAnnotID),
               ECMp=unique(ECMp), ECMpAnnotID=unique(ECMpAnnotID),
               MEP=unique(MEP),
               Row=unique(Row), Column = unique(Column), Block=unique(Block), ID=unique(ID),
               ArrayRow=unique(ArrayRow), ArrayColumn=unique(ArrayColumn),
               ArrayPrintOrder=unique(ArrayPrintOrder),ECMSourcePlateSeries=unique(ECMSourcePlateSeries),
               SCCmel = as.numeric(median(Spot_PA_SpotCellCount, na.rm=TRUE))),
         by="Barcode,Well,Spot"]

l3 <- l3[, SCCRel := calcResidual(SCCmel), by="Barcode,Well,MEP"]

```

```{r createMatrices}
#For an array  Unit of study there are 64 'samples'of 694 spots with 8 replicates
#Create a spot assignment that recognizes rotated B row arrays
l3$RSpot <- l3$Spot
l3$RSpot[grepl("B",l3$Well)] <- 701-l3$RSpot[grepl("B",l3$Well)]
#Add in ligand and ECMp names so they will carry through the normalization
l3$BWL <- paste(l3$Barcode,l3$Well,l3$Ligand,sep="_")
l3$SE <- paste(l3$RSpot,l3$ECMp,sep="_")
l3$ES <- paste(l3$ECMp,l3$RSpot,sep="_")
l3$BW <- paste(l3$Barcode,l3$Well,sep="_")
l3$WSE <- paste(l3$Well, l3$RSpot,l3$ECMp,sep="_")

```


```{r addResiduals}
#Cast to get mel values by barcode_well_ligand rows and spot columns
#Coerce missing values to have near 0 values
fill <- 0
l3Relc <- data.table(dcast(l3, BWL+ArrayPrintOrder+ECMSourcePlateSeries~ES, value.var=c("SCCRel"), fill = fill))

l3melc <- data.table(dcast(l3, BWL+ArrayPrintOrder+ECMSourcePlateSeries~ES, value.var=c("SCCmel"), fill = fill))

```


###Analysis of Spot Cell Count over Time

Analysis of Spot Cell Count over Time uses the raw spot cell counts and  residuals to assess the contents of the wells in the ECM source plates. Each set of ECM source plates is used to print a series of 8 well plates. The version 1 dataset in this analysis comes from three ECM source plate series 1, 3 and 4. 

The analysis method is:   
* Median summarize the cell level data to the spot level.  
* Create residuals by subtracting the median value of the replicates within each array from each spot value.  
* Create a matrix of residuals that is `r nrow(l3Relc)` rows (the number of arrays) by `r ncol(l3Relc)` columns (the number of spots in each array). Each column holds residuals for one spot in all MEMAs.  
* Create another matrix of the raw spot cell count values that is `r nrow(l3Relc)` rows by `r ncol(l3Relc)` columns, orgainzed similar to the residual matrix.  
* Plot a time course for each spot

```{r SpotTimeCourse}
setkey(l3Relc, BWL)
l3Relmelt <- melt(l3Relc, id.vars=c("BWL","ArrayPrintOrder","ECMSourcePlateSeries"), measure=patterns("_"),variable.name = "ES", value.name = paste0(unique(x[["Signal"]]),"Rel"))
# p <- ggplot(l3Relmelt, aes(x=PrintOrder, y=SCCRel1, colour=ES))+
#   geom_smooth(size=.1, se = FALSE)+
#   ggtitle("Time Course of Spot Cell Count Residuals\nColored by Spot")+
#   geom_vline(xintercept = c(seq(64,64*8,by=64)), color="blue",linetype = "dotdash", size=.1)+
#   guides(colour=FALSE)
# print(p)

p <- ggplot(l3Relmelt, aes(x=ArrayPrintOrder, y=SCCRel1, colour=ES))+
  geom_smooth(size=.1, se = FALSE)+
  ggtitle("Time Course of Spot Cell Count Residuals\nColored by Spot")+
  #geom_vline(xintercept = c(seq(64,64*6,by=64)), color="blue",linetype = "dotdash", size=.1)+
  facet_wrap(~ECMSourcePlateSeries,ncol=1, labeller = label_both)+
  guides(colour=FALSE)
print(p)
```

The plot below is interactive and cen be used to zoom in and identify indivdual spot data. The number after the underscore is the spot index in the array.  

<br>  

```{r RelPlotly}

ggplotly(p)

```

```{r SCCSpotTimeCourse}
setkey(l3melc, BWL)
l3melmelt <- melt(l3melc, id.vars=c("BWL","ArrayPrintOrder","ECMSourcePlateSeries"), measure=patterns("_"),variable.name = "ES", value.name = paste0(unique(x[["Signal"]]),"mel"))

p <- ggplot(l3melmelt, aes(x=ArrayPrintOrder, y=SCCmel1, colour=ES))+
  geom_smooth(size=.1, se = FALSE)+
  ggtitle("Time Course of Spot Cell Count \nColored by Spot")+
  #geom_vline(xintercept = c(seq(64,64*6,by=64)), color="blue",linetype = "dotdash", size=.1)+
  facet_wrap(~ECMSourcePlateSeries,ncol=1, labeller = label_both)+
  guides(colour=FALSE)
print(p)
```

The plot below is interactive and cen be used to zoom in and identify indivdual spot data. The number after the underscore is the spot index in the array.  

<br>  

```{r melPlotly}

ggplotly(p)

```
