---
output: html_document
---
---
title: "MEP-LINCS `r x[["CellLine"]]` `r x[["StainingSet"]]` `r x[["Signal"]]`  SVD Analysis"
date: "`r Sys.Date()`"
---

```{r global_options_Setup, include=FALSE}
#Author: Mark Dane, copyright 2015
knitr::opts_chunk$set(fig.width=12, fig.height=8,
                      echo=FALSE, warning=FALSE, message=FALSE)

#source("MEPLINCSFunctions.R")

library("ggplot2")
library("reshape2")
library("data.table")
library("limma")
library("MEMA")
library("grid")
library("knitr")
library("gplots")
library("RColorBrewer")
library("RUVnormalize")
library(plotly)

#Read in file based on name in dataframe
l1 <- fread(x[["inputFileName"]], showProgress = FALSE)

#Keep the control wells separate
l1$Ligand[l1$Ligand=="FBS"] <- paste(l1$Ligand[l1$Ligand=="FBS"],"C",as.numeric(factor(l1$Barcode[l1$Ligand=="FBS"])),sep="")
l1$MEP <- paste(l1$ECMp,l1$Ligand,sep="_")

#Delete the fiducial and blank spots
setkey(l1, "ECMp")
l1 <- l1[!"fiducial"]

#Create spot level dataset and 
#add Rel, the residual from subtracting the mel from each value

l3 <- l1[,list(Ligand=unique(Ligand), LigandAnnotID=unique(LigandAnnotID),
               ECMp=unique(ECMp), ECMpAnnotID=unique(ECMpAnnotID),
               MEP=unique(MEP),
               Row=unique(Row), Column = unique(Column), Block=unique(Block), ID=unique(ID),
               ArrayRow=unique(ArrayRow), ArrayColumn=unique(ArrayColumn),
               SCCmel = as.numeric(median(Spot_PA_SpotCellCount, na.rm=TRUE))),
         by="Barcode,Well,Spot"]

l3 <- l3[, SCCRel := calcResidual(SCCmel), by="MEP"]

```

```{r createMatrices}
#For an array  Unit of study there are 64 'samples'of 694 spots with 8 replicates
#Create a spot assignment that recognizes rotated B row arrays
l3$RSpot <- l3$Spot
l3$RSpot[grepl("B",l3$Well)] <- 701-l3$RSpot[grepl("B",l3$Well)]
#Add in ligand and ECMp names so they will carry through the normalization
l3$BWL <- paste(l3$Barcode,l3$Well,l3$Ligand,sep="_")
l3$SE <- paste(l3$RSpot,l3$ECMp,sep="_")
l3$ES <- paste(l3$ECMp,l3$RSpot,sep="_")
l3$BW <- paste(l3$Barcode,l3$Well,sep="_")
l3$WSE <- paste(l3$Well, l3$RSpot,l3$ECMp,sep="_")



```


```{r addResiduals}
#Cast to get mel values by barcode_well_ligand rows and spot columns
#Coerce missing values to have near 0 values
fill <- 0
l3Relc <- data.table(dcast(l3, BWL~ES, value.var=paste0(x[["Signal"]],"Rel"), fill = fill))

```


###Analysis of Spot Level Signals over Time

Analysis of Spot Level Signals over Time uses the residuals at the spot level as follows.  
* Median summarize the cell level data to the spot level.  
* Create residuals by subtracting the median value of the replicates from each spot value.  
* Create an M matrix of residuals that is `r nrow(l3Relc)` rows by `r ncol(l3Relc)` columns. Each column holds residuals for one spot in all MEMAs.  
* Plot a time course for each spot

```{r SpotTimeCourse}
setkey(l3Relc, BWL)
l3Relc$PrintOrder <- 1:nrow(l3Relc)
l3Relmelt <- melt(l3Relc, id.vars=c("BWL","PrintOrder"), measure=patterns("_"),variable.name = "ES", value.name = paste0(x[["Signal"]],"Rel"))
p <- ggplot(l3Relmelt, aes(x=PrintOrder, y=SCCRel1, colour=ES))+
  geom_smooth(size=.1, se = FALSE)+
  guides(colour=FALSE)
print(p)


```