---
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8,
                      echo=FALSE, warning=FALSE, message=FALSE)
```


```{r, echo=FALSE}
#Author: Mark Dane, copyright 2015

plotLEHmap <- function(dt, fill, titleExpression){
  p <- ggplot(dt, aes_string(x="Ligand", y="ECMp", fill = fill))+
    geom_tile()+
    scale_fill_gradient(low="white",high="red",oob = scales::squish,
                        limits=(quantile(dt[[fill]], probs=c(.01, .99), na.rm=TRUE)))+
    ggtitle(titleExpression)+
    xlab("")+ylab("")+
    guides(fill=FALSE)+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.6)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(.5)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.7)),panel.grid.major = element_line(linetype = 'blank'),panel.background = element_rect(fill = "dimgray"))
  suppressWarnings(print(p))
}


medianRUVIIIPlate <- function(k, Y, M, cIdx){
  #browser()
  nY <- RUVIII(Y, M, cIdx, k)
  #melt matrix to have ECMp and Ligand columns
  nYm <- melt(nY, varnames=c("Barcode","WSE"))
  #Extract ECMp and Well values into separate columns
  nYm$ECMp <- sub("([[:alnum:]]*_){2}","",nYm$WSE)
  nYm$Well <- strsplit2(nYm$WSE,split="_")[,1]

  #Median summarize by MEP
  nYm <- data.table(nYm)
  nYm <- nYm[, WSE := NULL]
  nYm <- nYm[,list(Value = median(value)), by="Barcode,Well,ECMp"]
  return(nYm)
}

madRUVIIIPlate <- function(k, Y, M, cIdx){
  #browser()
  nY <- RUVIII(Y, M, cIdx, k)
  #melt matrix to have ECMp and Ligand columns
  nYm <- melt(nY, varnames=c("Barcode","WSE"))
  #Extract ECMp and Well values into separate columns
  nYm$ECMp <- sub("([[:alnum:]]*_){2}","",nYm$WSE)
  nYm$Well <- strsplit2(nYm$WSE,split="_")[,1]

  #MAD summarize the normalized values by MEP
  nYm <- data.table(nYm)
  nYm <- nYm[, WSE := NULL]
  nYM <- nYm[,list(Value = mad(value)), by="Barcode,Well,ECMp"]
  return(nYM)
}

addMarginValues <- function(dt, mValue, MValue){
  #browser()
  dt.l <- dt[,list(m.l = median(get(mValue), na.rm = TRUE),
                   M.l = median(get(MValue), na.rm= TRUE)),
             by="Ligand"]
  
  dte. <- dt[,list(me. = median(get(mValue), na.rm = TRUE),
                   Me. = median(get(MValue), na.rm= TRUE)),
             by="ECMp"]
  
  setkey(dt,Ligand)
  setkey(dt.l,Ligand)
  dtm <- merge(dt,dt.l)
  
  setkey(dtm,ECMp)
  setkey(dte.,ECMp)
  dtmm <- merge(dtm, dte.) 
  return(dtmm)
}

plotCenteredBoxes <- function(dt, value, groupBy, colourBy=NULL, titleExpression){
  
  if(is.null(colourBy))  p <- ggplot(dt,aes_string(x=groupBy, y=value))
  else p <- ggplot(dt,aes_string(x=groupBy, y=value, colour=colourBy))
  
  p <- p + geom_boxplot()+ 
    ggtitle(titleExpression)+
    xlab("")+ylab("")+
    guides(fill=FALSE)+
    theme_grey(base_size = 12, base_family = "")+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(1)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(1)))
  suppressWarnings(print(p))
}

source("MEPLINCSFunctions.R")
supportingPlots = FALSE
runRUV2S = TRUE


```



```{r setup}
library("ggplot2")
library("reshape2")
library("data.table")
library("limma")
library("MEMA")
library("grid")
library("knitr")
library("gplots")
library("RColorBrewer")
library("RUVnormalize")

#Setup colors for Barcode and text in all heatmaps
selDark2 <- colorRampPalette(brewer.pal(8,"Dark2"))
plateCol = selDark2(8)

#ToDo Make changes for new folder sturcture
l1 <- fread(x[["inputFileName"]], showProgress = FALSE)

#Keep the control wells separate
l1$Ligand[l1$Ligand=="FBS"] <- paste(l1$Ligand[l1$Ligand=="FBS"],"C",as.numeric(factor(l1$Barcode[l1$Ligand=="FBS"])),sep="")
l1$MEP <- paste(l1$ECMp,l1$Ligand,sep="_")
setkey(l1, "ECMp")
l1 <- l1[!"fiducial"]
l1 <- l1[!"blank"]

#Logit transform DNA2NProportion
#logit(p) = log[p/(1-p)]
DNA2NImpute <- l1$Nuclei_PA_Cycle_DNA2NProportion
DNA2NImpute[DNA2NImpute==0] <- .01
DNA2NImpute[DNA2NImpute==1] <- .99
l1$Nuclei_PA_Cycle_DNA2NProportionLogit <- log2(DNA2NImpute/(1-DNA2NImpute))

#Create spot level dataset
l3 <- l1[,list(Ligand=unique(Ligand), LigandAnnotID=unique(LigandAnnotID),
               ECMp=unique(ECMp), ECMpAnnotID=unique(ECMpAnnotID),
               MEP=unique(MEP),
               Row=unique(Row), Column = unique(Column), Block=unique(Block), ID=unique(ID),
               ArrayRow=unique(ArrayRow), ArrayColumn=unique(ArrayColumn),
               EdUmel = median(Nuclei_PA_Gated_EduPositiveLogit, na.rm=TRUE),
               DNA2Nmel = median(Nuclei_PA_Cycle_DNA2NProportionLogit, na.rm=TRUE),
               SCCmel = as.numeric(median(Spot_PA_SpotCellCount, na.rm=TRUE))),
         by="Barcode,Well,Spot"]


l1MEPs <- l1[,list(mel = median(Nuclei_PA_Gated_EduPositiveLogit, na.rm = TRUE),
                   Mel = mad(Nuclei_PA_Gated_EduPositiveLogit, na.rm= TRUE)),
             by="Barcode,Well,Ligand,ECMp,MEP"]

#l1MEPS is the data.table of the MEP level, raw, logit-transformed EdU responses
l1MEPs <- addMarginValues(l1MEPs,"mel","Mel")

#calculate the overall median and MAD for the staining set
m.. <- median(l1MEPs$mel, na.rm = TRUE)
M.. <- median(l1MEPs$Mel, na.rm = TRUE)
```

```{r RUVIII, eval=TRUE}
#Order the spot level data by plate, well, spot number
#Create a spot assignment that works with rotated B row arrays
#RUV-III without the final step of averaging. 
#k=0,1, 2, 3, 4 say
#treat plates as units 
#neg controls from control ligand
#view all 8 plates as replicates.
#M would be a m by 1 matrix of all 1s, where m is the number of rows of Y.  
#keep k very small.

l3$RSpot <- l3$Spot
l3$RSpot[grepl("B",l3$Well)] <- 701-l3$RSpot[grepl("B",l3$Well)]
#Add in ligand and ECMp names so they will carry through the normalization
l3$BWL <- paste(l3$Barcode,l3$Well,l3$Ligand,sep="_")
l3$SE <- paste(l3$RSpot,l3$ECMp,sep="_")
l3$BW <- paste(l3$Barcode,l3$Well,sep="_")
#Cast to get mel values by barcode_well rows and spot columns
#Coerce missing values to have near 0 proliferation signals
l3EdUc <- dcast(l3, BWL~SE, value.var="EdUmel",fill = log2(.01/(1-.01)))
#Remove the BWL column and use it as rownames in the matrix
l3EdUm <- l3EdUc[,grep("BWL",colnames(l3EdUc), value=TRUE, invert=TRUE), with=FALSE]
#YEdU is a numeric matrix of the raw logit transformed EdU responses 
#of each spot extracted from the l3 dataset
YEdU <- matrix(unlist(l3EdUm), nrow=nrow(l3EdUm), dimnames=list(l3EdUc$BWL, colnames(l3EdUm)))

```
---
title: "MEP-LINCS `r x[["CellLine"]]` `r x[["StainingSet"]]` `r x[["Signal"]]` `r x[["Method"]]` Normalization"
date: "`r Sys.Date()`"
---

###Proliferation Signal
This analysis starts with a focus on the proportion of EdU positive cells at each spot.
For each plate, k-means clustering (k = 2) is used on the raw EdU median intensity values of the cells in the A03 control well. Cells in the higher-value cluster are labeled as EdU+.  
The intensity value that divides the control well clusters is used as a threshold to label the cells as EdU+ in the seven ligand wells of each plate.  
The proportion of EdU+ cells at each spot is calculated.  
The proportion is transformed by the logit function defined as logit(p) = log2(p/1-p) which takes on values from negative to positive infinity. The proportions in this dataset have been transformed into the range [.01, .99] to eliminate issues with logit transformed values that are infinite.  


```{r, eval=TRUE, fig.width=4.5, fig.height=4}


titleExpression <- expression(EdU^{"+"}~ Medians~(m["e,l"]))
plotLEHmap(l1MEPs, fill="mel", titleExpression)

titleExpression <- expression(EdU^{"+"}~MADs~(M["e,l"]))
plotLEHmap(l1MEPs, fill="Mel", titleExpression)


```

###RUVIII with Plate as Unit of Study

In the following plots RUV-3 normalization is applied to the EdU signal of all spots with the plate as the unit of study. This creates an 8x5552 matrix of EdU+ spot values. The 694 spots in the A03 control wells of each plate are designated as negative controls. The faceted plots show normalizations at various k values.

```{r}
#Setup data with plate as the unit
#There are 8 'samples' with 694 controls
#Order the spot level data by plate, well, spot number
#Add in ligand and ECMp names so they will carry through the normalization
l3$WSE <- paste(l3$Well, l3$RSpot,l3$ECMp,sep="_")
#Cast to get mel values by barcode by well+Spot+ECMp
#Use the names to hold the spot contents
#Coerce missing values to have near 0 proliferation signals
l3EdUPlateC <- dcast(l3, Barcode~WSE, value.var="EdUmel",fill = log2(.01/(1-.01)))
#Remove the Barcode column and use it as rownames in the matrix
l3EdUPlateM <- l3EdUPlateC[,grep("Barcode",colnames(l3EdUPlateC), value=TRUE, invert=TRUE), with=FALSE]
YEdUPlate <- matrix(unlist(l3EdUPlateM), nrow=nrow(l3EdUPlateM), dimnames=list(l3EdUPlateC$Barcode, colnames(l3EdUPlateM)))

```

```{r}

nYEdUPlateList <- lapply(1:4, function(k, Y, M, cIdx, mValue, MValue){
    #browser()
    nYEdUm <- medianRUVIIIPlate(k, Y, M, cIdx)
    setnames(nYEdUm, "Value", "mel")
    nYEdUM <- madRUVIIIPlate(k, Y, M, cIdx)
    setnames(nYEdUM, "Value", "Mel")
    setkey(nYEdUm,Barcode,Well,ECMp)
    setkey(nYEdUM,Barcode,Well,ECMp)
    nYEdU <- merge(nYEdUm, nYEdUM)
    nYEdU$BW <- paste(nYEdU$Barcode,nYEdU$Well, sep="_")
    setkey(nYEdU,BW)
    BWL <- unique(l3[,.(BW,Ligand)])
    setkey(BWL,BW)
    nYEdU <- merge(nYEdU, BWL)
    nYEdU$MEP <- paste(nYEdU$ECMp,nYEdU$Ligand, sep="_")
    nYEdU <- nYEdU[,BW :=NULL]

    #Add margin values
    nYEdU <- addMarginValues(dt=nYEdU, mValue="mel",MValue="Mel")
    nYEdU$k <- k
    return(nYEdU)
}, Y=YEdUPlate, M=matrix(1,nrow=nrow(YEdUPlate)), cIdx=which(grepl("A03",colnames(YEdUPlate))), mValue="mel", MValue="Mel")

nYEdUPlatek <- rbindlist(nYEdUPlateList)

```


```{r, fig.height=3}
groupBy <- "ECMp"
value <- "mel-m.l"
colourBy <- "Barcode"
titleExpression=expression(EdU^{"+"}~Centered~Medians~by~ECMp~(m["e,l"]*"-"*m[".,l"]))
p <- ggplot(l1MEPs, aes_string(x=groupBy, y=value))+
  geom_boxplot()+ 
  ggtitle(titleExpression)+
  xlab("")+ylab("")+
  ylim(-6,4)+
  guides(fill=FALSE)+
  theme_grey(base_size = 12, base_family = "")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.6)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(.5)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.5)))
suppressWarnings(print(p))
```

```{r}

dt <- nYEdUPlatek

p <- ggplot(dt,aes_string(x=groupBy, y=value))+
  geom_boxplot()+ 
  ggtitle(bquote(EdU^{"+"}~ Medians~(m["e,l"]*"-"*m[".,l"])~After~RUVIII~Normalization~with ~Plate~Unit~of~Study))+
  xlab("")+ylab("")+
  ylim(-6,4)+
  guides(fill=FALSE)+
  theme_grey(base_size = 12, base_family = "")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.6)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.5)))+
facet_grid(k~., labeller = labeller(k = label_both))
suppressWarnings(print(p))
```

```{r, fig.height=3}
groupBy <- "Ligand"
value <- "mel-me."
colourBy <- "Barcode"
titleExpression=expression(EdU^{"+"}~Centered~Medians~by~Ligand~(m["e,l"]*"-"*m["e,."]))
p <- ggplot(l1MEPs, aes_string(x=groupBy, y=value, colour=colourBy))+
  geom_boxplot()+ 
  ggtitle(titleExpression)+
  xlab("")+ylab("")+
  ylim(-6,4)+
  guides(fill=FALSE)+
  theme_grey(base_size = 12, base_family = "")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.6)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(.5)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.5)))
suppressWarnings(print(p))

```

```{r}

dt <- nYEdUPlatek

p <- ggplot(dt,aes_string(x=groupBy, y=value, colour=colourBy))+
  geom_boxplot()+ 
  ggtitle(bquote(EdU^{"+"}~ Medians~(m["e,l"]*"-"*m["e.,."])~After~RUVIII~Normalization~with ~Plate~Unit~of~Study))+
  xlab("")+ylab("")+
  ylim(-6,4)+
  guides(fill=FALSE)+
  theme_grey(base_size = 12, base_family = "")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.6)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.5)))+
  facet_grid(k~., labeller = labeller(k = label_both))
suppressWarnings(print(p))

```


```{r, fig.width=5, fig.height=4}

fill <- "mel"
titleExpression <- expression(EdU^{"+"}~ Medians~(m["e,l"]))
dt <- l1MEPs
limits <- quantile(dt[[fill]], probs=c(.01, .99))
                 
p <- ggplot(dt, aes_string(x="Ligand", y="ECMp", fill = fill))+
  geom_tile()+
  scale_fill_gradient(low="white",high="red",oob = scales::squish,limits=quantile(dt[[fill]], probs=c(.01, .99)))+
  ggtitle(titleExpression)+
  xlab("")+ylab("")+
  #guides(fill=FALSE)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.6)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(.5)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)),panel.grid.major = element_line(linetype = 'blank'),panel.background = element_rect(fill = "dimgray"))
suppressWarnings(print(p))
```

```{r, fig.width=4.5}

dt <- nYEdUPlatek

p <- ggplot(dt, aes_string(x="Ligand", y="ECMp", fill = fill))+
  geom_tile()+
  scale_fill_gradient(low="white",high="red",oob = scales::squish, limits=quantile(dt[[fill]], probs=c(.01, .99)))+
   ggtitle((bquote(EdU^{"+"}~ Medians~(m["e,l"]*"-"*m["e.,."])~After~RUVIII~Normalization~with ~Plate~Unit~of~Study)))+
  facet_grid(k~., labeller = labeller(k = label_both))+
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.6)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(.4)), plot.title = element_text(size = rel(.7)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)),panel.grid.major = element_line(linetype = 'blank'),panel.background = element_rect(fill = "dimgray"))
suppressWarnings(print(p))

```


```{r, fig.width=5, fig.height=4}

fill <- "Mel"
titleExpression <- expression(EdU^{"+"}~ MADs~(M["e,l"]))
dt <- l1MEPs
limits <- quantile(dt[[fill]], probs=c(.01, .99))
                 
p <- ggplot(dt, aes_string(x="Ligand", y="ECMp", fill = fill))+
  geom_tile()+
  scale_fill_gradient(low="white",high="red",oob = scales::squish,limits=quantile(dt[[fill]], probs=c(.01, .99)))+
  ggtitle(titleExpression)+
  xlab("")+ylab("")+
  #guides(fill=FALSE)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.6)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(.5)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)),panel.grid.major = element_line(linetype = 'blank'),panel.background = element_rect(fill = "dimgray"))
suppressWarnings(print(p))
```

```{r, fig.width=4.5}

dt <- nYEdUPlatek

p <- ggplot(dt, aes_string(x="Ligand", y="ECMp", fill = fill))+
  geom_tile()+
  scale_fill_gradient(low="white",high="red",oob = scales::squish, limits=quantile(dt[[fill]], probs=c(.01, .99)))+
   ggtitle((bquote(EdU^{"+"}~ MADs~(M["e,l"]*"-"*M["e.,."])~After~RUVIII~Normalization~with ~Plate~Unit~of~Study)))+
  facet_grid(k~., labeller = labeller(k = label_both))+
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.6)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(.4)), plot.title = element_text(size = rel(.7)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)),panel.grid.major = element_line(linetype = 'blank'),panel.background = element_rect(fill = "dimgray"))
suppressWarnings(print(p))

```
